---
title: "Oct 1 Activities: M&Ms; Survey; Sampling"
date: 9-30-2025
---

All three class activities are on this .rmd notebook

Activity #1: Activity: M&Ms - Central Limit Theorem

Working in groups, each group will get a handful of candy from the bag, and then count the number of red and black candies. Log those in this Google Sheet 

https://docs.google.com/spreadsheets/d/1KOeoVTq2W7L2T9sWjwh5U_2f59GsTeAylVGD53rNWww/edit?usp=sharing

(give your group a nice name - Outta the park).

Take a look at the values that we all get - what do you notice about the patterns? This is an exercise about the central limit theorem. 

The central limit theorem states that the sampling distribution of the mean of any independent, random variable will be normal or nearly normal, if the sample size is large enough.

Then discuss and try to answer these questions:

--How close is your group's mean to the true proportion of 0.5?
--Why do different groups get different results?
--What happens to individual extreme results when we combine all groups?
--How does this relate to polls and surveys in real life?

Write your answer below. And then you may eat the M&Ms. 

**A:**

For most groups, the percentage of brown MnMs fell around 90% (+/- 5%). Different groups may get different results because its a random sample. Some groups have a higher total, some have less. If we combine all groups, the extreme results effect the average, bringing it up or down. In real life, sample size isn't always representative of the population. 

# Activity #2:

Belonging Survey

Working in pairs, take a look back at the main findings from UMD's Belonging Survey:

https://president.umd.edu/internal/belonging


Specifically, focus on the "Sample Description" section and the Appendix. Discuss the differences and between the population and the sample, and what the document says about how the survey attempted to deal with them. Are there any decisions described in the document that raise questions in your minds? Detail the differences you see and your understanding of the methods they employed (it's ok if you don't understand all of them), and add your questions. You should discuss this as a pair but submit individually.

**A:**

We thought it was interesting that the majority of responses came from UMD staff and faculty, despite the fact that overall, students make up the majority of the UMD population. Therefore, this doesn't accurately represent the feelings of the whole UMD population, it skews the results to reflect more of the staff sentiment.
In the Appendix, I thought it was particularly interesting how there weren't enough Jewish respondant for individual analysis in the survey, despite the fact the roughly 1/5 of the UMD population identifies as Jewish. Overall, the survey responses did not represent the true nature of UMD's population.

# TASK 3: SAMPLING

Based on the class sampling demo (available on Elms), sample data from a population using R and the Tidyverse. You should get started by loading the tidyverse in a new R block.

### Task 1: Load the tidyverse

**A**
```{r}

library(tidyverse)

```


### Task 2: Load 1,000 rows of data with the ages of people from 18-80

**A**
```{r}
set.seed(42)
pop_ages <- tibble(
  age = sample(18:80, 1000, replace = TRUE)
)

head(pop_ages)

```

### Task 3: Sample 100 rows of data from the dataset
**A**
```{r}
sample_100 <-pop_ages |> 
  sample_n(100)

sample_75 <-pop_ages |> 
  sample_n(75)

sample_50 <-pop_ages |> 
  sample_n(50)

sample_25 <-pop_ages |> 
  sample_n(25)

```


### Task 4: Calculate the mean age and standard deviation of the sample
**A**
```{r}
sample_100 |> summarize(mean = mean(age), median = median(age), sd = sd(age))
sample_75 |> summarize(mean = mean(age), median = median(age), sd = sd(age))
sample_50 |> summarize(mean = mean(age), median = median(age), sd = sd(age))
sample_25 |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 5: Calculate the mean age and standard deviation of the population
**A**
```{r}
pop_ages |> summarize(mean = mean(age), median = median(age), sd = sd(age))

```


### Task 6: Compare the sample and population statistics

The sample median and mean were slightly lower than the population ones. 

Compare the means and standard deviations. Are they similar? Go back and re-run the sampling and statistics calculations with a smaller sample size. How do the results change?

The means and medians fluctuated when I went back and tried different sample sizes. However, they are all similar, typically within a value of +/- 5.

When finished, save your work, switch to GitHub Desktop, then add, commit and push your changes to GitHub and submit the URL of the notebook in ELMS.
