---
title: "midterm_key_oct_8"
author: "Rob Wells"
date: "2025-10-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# JOUR405: Statistics for Journalists
## Midterm Exam - Fall 2025

For this exam, you'll analyze several datasets using R and the statistical concepts we've covered in class. Load the tidyverse before beginning, then complete each task. Write your code in the provided blocks and answer the questions in complete sentences.

# Load tidyverse
```{r}
library(tidyverse)
```

## Part 1: Restaurant Health Inspections (25 points)

The first dataset contains restaurant health inspection scores from Montgomery County. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv`

The dataset includes the name of the establishment, the number of points for critical and non-critical areas, the total points, maximum points possible and the compliance score and grade.

## Task 1: Read the data**
```{r}
health_inspections <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/montco_inspections.csv")
```


## Task 2: Calculate the mean and standard deviation maximum and minimum of compliance scores. Put the results in a new dataframe called "compliance_stats." 
```{r}
compliance_stats <- health_inspections |>
  summarize(mean_score = mean(compliance_score),
              sd_score = sd(compliance_score),
              max_score = max(compliance_score),
              min_score = min(compliance_score)
            )

```

## Task 3: Create a histogram of the compliance scores with a vertical line showing the mean. Include a headline. Include a headline and label the x and y axes.
```{r}
health_inspections |>
    ggplot() + 
  geom_histogram(aes(x = compliance_score), binwidth = 10) +
  geom_vline(aes(xintercept = mean(compliance_score)), color = "red", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(compliance_score) - sd(compliance_score)), color = "blue", linetype = "dashed", size = 1) +
  geom_vline(aes(xintercept = mean(compliance_score) + sd(compliance_score)), color = "blue", linetype = "dashed", size = 1) +
  
  labs(title = "Compliance Scores for Restaurants",
       x = "Compliance Score",
       y = "Number of Restaurants")




```

## Task 4: Write 2-3 sentences interpreting what the standard deviation and histogram tell us about the distribution of compliance scores. What would be newsworthy about this distribution? What's the story here? 

The histogram of the compliance scores is skewed to the left. The first standard deviation (as seen in the histogram) tells us that over 68% of restaurants scored above a 90 on their compliance scores. In fact, the average was 96.3. What's newsworthy about this distribution is that more than half of restaurants have a score over 90 for compliance. The majority of restaurants in Montgomery county are doing relatively well in terms of their complience. It would be interesting to see how this compares to other counties.
Additionally, what standards does the county hold their restaurants too? Is there a reason they are generally doing pretty well, possibly compared to other counties?


## Part 2: High School Athletics (30 points)

The second dataset shows participation numbers in high school sports across Maryland counties in 2024, broken down by sport and gender. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv`

# Task 5: Load the high school participation data
```{r}
sports_participation <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_hs_participation.csv")
```


#Task 6: Add two columns called total and girls_pct.  
Add two columns called total and girls_pct using mutate(), with the total adding together boys and girls and girls_pct being the percentage of the total represented by girls participants. 
```{r}
sports_participation <- sports_participation |>
  mutate(total = rowSums(across(3:4)),
         girls_pct = ((girls *100)/ total)
           )
```

#Task 7: Calculate correlation between boys and girls participation
```{r}
sports_participation |>
  summarize(correlation = cor(boys, girls, method="pearson"))

```

#Task 8: Create a scatterplot showing this relationship, adding a line of best fit. Include a headline and label the x and y axes.
```{r}
sports_participation |>
 ggplot() +
  geom_point(aes(x=boys, y=girls)) +
  geom_smooth(aes(x=boys, y=girls), method="lm")+
  expand_limits(x = 0, y = 0) +
   labs(title = "Correlation between boys and girls participation in sports",
       x = "Boys",
       y = "Girls")


```


# Task 9: In 2-3 sentences, explain what the correlation coefficient and scatterplot reveal about gender equity in Maryland high school sports participation. How do you interpret the school districts that are below the line vs those that are above? Which school districts are most worth examining further, and why? (10 points)

Overall, the participation of girls and boys in sports in Maryland high schools has a very high correlation. Generally, those over the line have more girls than average, while those below have more boys than average. It might be worth examining  what is different about those districts with more girls because the majority of schools fall slightly below the line. Those above also have a bit of a wider gap which is interesting.

## Part 3: Maryland Car Theft Rates (25 points)

Your editor has assigned you a story about car thefts in Maryland and wants you to analyze the data to find out which counties have the highest rates. The fourth dataset contains car theft statistics for Maryland counties in 2023 and population. Load the data from: `https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv` and complete the following tasks:

#Task 10: Load the car thefts data
```{r}
car_thefts <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_car_thefts.csv")
```

#Task 11: Using mutate, add a column that calculates the rate of car thefts for each county - you need to choose the per capita rate 
```{r}
car_thefts <- car_thefts |>
  mutate(rate_county = (X2023/population * 10000)
  )
```

#Task 12: Calculate the median car theft rate and the total number of car thefts statewide. Which counties have rates above the median? 
```{r}
car_thefts |>
  summarise(median_thefts = median(rate_county),
            theft_total = sum(X2023),
            correlation = cor(population, rate_county, method="pearson")
            )
```


#Task 13: Write 2-3 sentences describing what these calculations reveal about the distribution of car thefts in Maryland. What's the lede of a story about your findings? 

The distribution of car thefts in Maryland are a bit all over the place. The median theft rate was 9.5, however, the highest rate by far was in Baltimore County, which had a rate of 123.7, and the lowest was Somerset County with 2.1. I was curious to see if there was a correlation between population and the rate per county, but the the correlation wasn't the strongest.Therefore, what does Somerset County do that Baltimore doesn't? What other factors may attribute to the car theft rate?
Lede: Somerset County, Maryland saw the lowest car theft rate per capita in the whole state. How they address this issue compared to other counties in the state.

## Part 4: Data Analysis Scenario (20 points)

#Task 14: Essay Question.

You receive a tip that local emergency response times have gotten significantly worse over the past year. You obtain monthly data on response times for police, fire and ambulance calls.

Write 3-4 sentences explaining:
1. What statistical measures would you calculate to verify this claim? (10 points)
2. What visualizations would help readers understand the trends? (5 points)
3. What additional context or data would you need to make this a complete story? (5 points)

**Write your answer here:**

The statistical measures that I would address are the amount of calls to emergency response teams in the past year, how many people work at these call centers, and the average time people are waiting. I think the call volume and the number of workers may be significant factors in how long call times are, I would compare these measures.
I think a scatter plot would be most efficient in comparing these values, especially if there is a high correlation between less workers and longer wait times.
I think the hiring budget of different counties would be interesting additional data, perhaps they smaller budgets would mean less workers, resulting in longer wait times.


### Submission Instructions
- Save your work frequently
- Make sure all code blocks run without errors
- Provide clear explanations for your analytical choices
- Submit both your code and written responses
- Load this test to GitHub and submit the link in Elms

Good luck!
