---
title: "HW3 WH Salaries"
---

In this assignment, we will be working with a dataset containing information about the salaries of White House staffers as of July 2024.

### Task 1: Load the Tidyverse

You will need to load the tidyverse library

```{r}
library(tidyverse)

```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv. You will create a dataframe called `wh_salaries` that contains the data from this URL.

### Task 2: Load the Data and Describe It

```{r}
wh_salaries <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/biden_wh_salaries_2024.csv")
```
### Task 3: Describe the Data

You should examine that dataframe carefully by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the categories of incidents. From a reporting standpoint, what stands out to you about this data? Are there any salary values that will cause problems for analysis?

An interesting observation I has was that people who were special/deputy assistants (there are a lot!) to the resident or a part of a council general earned over 100,000. While others earned under. Advisors earned 0, even those that were catergorized as employees. I'd be interested, as a reporter, to know about the employees that are eemingly advising for free, and if they get anything out of it.

### Task 4: Calculate the Mean and Median of the Salaries

Write code to calculate the mean and median of the salaries in the dataset using the `summarize()` function. You don't need to make a new dataframe, just display the results.

```{r}
library(dplyr)

wh_salaries |>
  mutate(salary = as.numeric(gsub("[^0-9.]", "", salary))) |>
  summarize(
    mean_salaries = mean(salary, na.rm = TRUE),
    median_salaries = median(salary, na.rm = TRUE)
  )


```

### Task 5: Add in the minimum and maximum salaries

Take the code you just wrote and add in the minimum and maximum salaries (the functions for that are `min()` and `max()`). Your results should show four columns: mean_salary, median_salary, min_salary, max_salary.

```{r}
wh_salaries |>
  mutate(salary = as.numeric(gsub("[^0-9.]", "", salary))) |>
  summarize(
    mean_salaries = mean(salary),
    median_salaries = median(salary),
    max_salaries = max(salary),
            min_salaries = min(salary)
  )
```

### Task 6: Describe the Results

Write a couple of sentences describing the results of your analysis. What do the mean and median salaries tell you about the data? What do the minimum and maximum salaries tell you? Are there any outliers in the data?

The mean and median salaries are only 3,000 dollars apart, this tells me that the median is almost a true "middle" of the data. The average WH employee makes over 100,000 dollars. However, the maximum is almost double that, while the minimum is 0. It seems that the minimum before 0 is around 56,000, which is about half as the average/median. I wonder how the average/median would change without those 0 values.

### Task 7: Story Idea

Based on the data, what is your best story idea? What is the next step you would take to develop that story idea?

As I proposed earlier, I'd be curious to find out more about the advisers seemingly advising for free. They are considered employees, however, do they receive compensation in a different capacity? I think this could make an interesting story, especially regarding those considered employees. My next step would be to dive deeper into who exactly these people are and gather more info on their role in the administration. What makes them employees, compared to other advisers who "detailees." How much work do they usually do despite not being paid?
