---
title: "HW4 MD CityCrime"
---

In this assignment, we will be working with a dataset of total crimes reported by Maryland cities for two consecutive years. You'll load the data, calculate the violent and property crime rates for each city for both years, and then calculate the percentage change in the crime rates between the two years. You'll also write a short story based on your findings.

### Task 1: Load the Tidyverse and Janitor

You will need to load the tidyverse library and the janitor library

```{r}
library(tidyverse)
library(janitor)
```

Then you'll load the data from the web. The URL for the data is: https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv. You will create a dataframe called `md_cities` that contains the data from this URL.

### Task 2: Load the Data

```{r}
md_cities <- read.csv("https://raw.githubusercontent.com/dwillis/jour405/refs/heads/main/data/md_cities_crime_2019_2020.csv")
```
### Task 3: Calculating Rates

Calculate violent and property crime rates for each city for both years using `mutate()`. You'll create four new columns: `violent_rate_2019`, `property_rate_2019`, `violent_rate_2020`, and `property_rate_2020`. You'll need to figure out the proper unit rate, given that there are varying populations. There's no perfect choice; choose the one that works best for you. Save the results to a new dataframe called `md_cities_rates`.

```{r}
md_cities_rates <- md_cities |>
  mutate(
    violent_rate_2019 = (pop2019 / violent2019),
    property_rate_2019 = (pop2019 / property2019),
    violent_rate_2020 = (pop2020 / violent2020),
    property_rate_2020 = (pop2020 / property2020)
  )
```

### Task 4: Describe the Rates

Examine the new dataframe, `md_cities_rates`, by clicking on it in the Environment tab in RStudio and sorting by the rate columns. Write a couple of sentences about the results, comparing and contrasting the cities with the highest rates (both violent and property) in 2019 and 2020. Do you see any patterns? What does this tell you about crime in Maryland cities? What's the story here?

It seems like in general, violent crimes occur at a less frequent rate than property crimes. For countless cities, some experienced 0 crimes for certain years, making their rate for that type of crime non-existent. I'd be interested to see what those cities are doing to prevent crime compared to those with higher numbers. Or are there other reasons why these cities experience less crimes? Some cities, like Emmittsburg, Middletown, Walkersville, Mt. Airy, etc had a substantially larger crime rate than other cities, why is that? Do counties have anything to do with it?

### Task 5: Calculate Percentage Change

Let's add two percentage change columns to your dataframe: one for violent crime rates and one for property crime rates. You'll calculate the percentage change between 2019 and 2020 for each city. Save the results to a new dataframe called `md_cities_change`.

```{r}
md_cities_change <- md_cities_rates |>
  mutate(
    pct_violent_change = ((violent_rate_2020 - violent_rate_2019) / violent_rate_2019) * 100,
    pct_property_change = ((property_rate_2020 - property_rate_2019) / property_rate_2019) * 100
  )


```

### Task 7: Describe Percentage Changes

Look at the `md_cities_change` dataframe by clicking on it in the Environment tab in RStudio. Write a couple of sentences describing the percentage change results. What's more interesting here, the patterns or the outliers? What's the story in the data?

It seemed pretty common that cities with a decrease in the rate of one crime usually saw a decrease in the other type of crime. I found this interesting because it appears indicative of a change in local policy or citizen behavior. Of course, there are outliers with this. Its also interesting for cities that had no crime in both years, of course thier percent change was non-exitent.

### Task 8: Tell the Best Story

Write up a lede and nutgraf of a story with the _best_ story idea you've found in this assignment. The lede should be a single sentence that summarizes the most important information in the data. The nutgraf should be a paragraph that provides more context and explains why the data is important.

Lede: “Carroll County saw an almost across-the-board rise in crime between 2019 and 2020.”

Nutgraf: "The crime rate in Carroll County, Maryland rose in nearly every city between 2019 and 2020, in both violent and property offenses. That makes it the most consistent increase in the state, where other counties saw more mixed changes."

### Task 9: Save Your Work and Put it on GitHub

Save this notebook, then commit it to your local repository. Push the changes to your GitHub repository and submit the URL for this file in ELMS.
