---
title: "final_project"
author: "Kaya Bogot"
date: "2025-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)  # For data manipulation and visualization
library(lubridate)  # For handling dates
library(knitr)      # For nice tables
library(scales)     # For better axis formatting
```
```{r}
billboard <- read_csv("billboard_hot _100_number ones_database_data.csv")
data_dictionary <- read_csv("data_dictionary.csv")
lyrical_topics <- read_csv("lyrical_topics.csv")
```

-- What does it contain and what time period does it cover?

The billboard chart contains information about songs and their artists, along with data about their time on billboard charts, their genre, danceability, instruments, etc.
It covers the time period from 1958-2025.

--Which columns will be most important for your analysis?

Danceability, song structure, and weeks at number 1.

--What's missing that you might need or want?

Copies sold/streams, depending on how it was consumed at it's release. To determine general popularity besides chart numbers.

--Select a relevant column and produce basic descriptive statistics on the mean, median, lowest and highest values.

```{r}
billboard |> 
  summarize(
    mean_Danceability   = mean(Danceability, na.rm = TRUE),
    median_Danceability = median(Danceability, na.rm = TRUE),
    max_Danceability    = max(Danceability, na.rm = TRUE),
    min_Danceability    = min(Danceability, na.rm = TRUE),
    sd_Danceability = sd(Danceability, na.rm = TRUE)
  )

```

For this step, I just wanted to get a baseline of the data surrounding Danceability.

Next, propose a newsworthy hypothesis (or several of them) for your dataset. You'll be looking to find (or eliminate possible story ideas) using the techniques you've learned in this class. You'll start that effort this week by coming up with some theories or questions that you're interested in exploring that can be answered with the dataset you selected.

Null Hypothesis: Song structures do not effect the popularity of songs, nor does it effect the danceability.

Alternative Hypothesis: Specific song structures are more popular than others. Leading to more weeks at number 1. Song structure also effects danceability, with certain structures yielding higher danceability scores. 

--What questions do you have about it or need help answering?

Is there a simple way I can find streams/copies sold? Or will I have to search 1 by 1? I wasn't sure if there was a database or something.

**Part 2**

**Fix and revise your project based on the grading feedback.**

**Run at least two advanced statistical tests (t-test, Anova, z-score, etc.) that fit your project. You may need to revise the null and alternate hypothesis as necessary.**

```{r}
z_scores <- billboard |>
  transmute(

    z_weeks_number_one = (`Weeks at Number One` - mean(`Weeks at Number One`, na.rm = TRUE)) /
                          sd(`Weeks at Number One`, na.rm = TRUE),
    z_danceability     = (Danceability - mean(Danceability, na.rm = TRUE)) /
                          sd(Danceability, na.rm = TRUE)
  )


billboard |>
  select(`Weeks at Number One`, Danceability) |>
  bind_cols(z_scores) |>
  head(10)



```

For the z-scores, there were no obvious standouts. There were no scores significantly below -2 or above +2, which would be unusual. 

```{r}
anova_dance <- aov(Danceability ~ `Song Structure`, data = billboard)
summary(anova_dance)

anova_weeks <- aov(`Weeks at Number One` ~ `Song Structure`, data = billboard)
summary(anova_weeks)

```
The ANOVA results show that song structure is statistically significant for both danceability and popularity, allowing us to reject the null hypothesis in both cases. For danceability, the results produced an F-statistic of 4.44 and a p-value of 4.38e-15 (way below 0.05, meaning it is statistically significant), meaning that certain song structures are more indicative of danceability that it would be by chance. Therefore, there is a strong relationship between song structure and how danceable a song is.

For popularity, the F-statistic was 1.52 and the p-value was 0.0306, which is statistically significant but less so than danceability. This means that while song structure influences popularity, it explains less of the overall results than danceability.


```{r}

billboard_summary <- billboard|>
  group_by(`Song Structure`) |>
  summarise(
    mean_danceability = mean(Danceability, na.rm = TRUE),
    mean_weeks_number_one = mean(`Weeks at Number One`, na.rm = TRUE),
    n_songs = n()
  ) |>
  arrange(desc(n_songs))  

billboard_summary
```

This plot was just to organize the data and see how many of each stong structure these was. Plus, it calculated the mean danceability and popularity of each structure.

**Create at least one data visualization (use ggplot)**

```{r}
structure_popularity <- billboard

ggplot(structure_popularity) +
  geom_point(aes(x = `Song Structure`, y = `Weeks at Number One`)) +
  geom_smooth(aes(x = `Song Structure`, y = `Weeks at Number One`), method = "lm") +
  expand_limits(x = 0, y = 0)

```
This graph is interesting because there appears to be a bit of a skew to the right. This isn't exactly a histogram, but it vaguely takes the shape of one. We can see that more songs in the "C" and "E" category of song structures spent more time at number one on the billboard chart.This is mostly apparent when looking above the 10 week mark. These findings reinforce my alternative hypothesis that specific song structures are more popular than others, leading to more weeks at number 1.

**Conclusion**

Over the course of this project, I have analyzed how song structure effects the danceability and popularity of songs. By alternative hypothesis being that "specific song structures are more popular than others, leading to more weeks at number 1. Song structure also effects danceability, with certain structures yielding higher danceability scores." By first running an ANOVA test, it was clear that both factors were statistically significant in predicting the danceability and popularity of a song. Additionally, by creating a visualization of the relationship between weeks at number one and song sturctures, I was able to visualize exactly which categories yielded more weeks at number 1. "C" and "E" structures generally had more songs at number 1 for more than 10 weeks.
As a reporter, it would be interesting to see if this kind of data is being used in the music industry to create hits. It would also be interesting to examine the psychology of why these songs structure are more appealing to listeners. In a story such as these, I would use this data and talk to experts in psychology and music about the "why" behind this correlation and find out more about the causation. I would also talk to listeners and find out their opinions on songs and why they think there is a causation. One potential porblem, however, is if this song structure system is universal, or if it is just unique to the creator of this data set. If it is not universal, it would require more research or cross-referencing to existing systems.

