---
title: "final_project"
author: "Kaya Bogot"
date: "2025-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)  # For data manipulation and visualization
library(lubridate)  # For handling dates
library(knitr)      # For nice tables
library(scales)     # For better axis formatting
```
```{r}
billboard <- read_csv("billboard_hot _100_number ones_database_data.csv")
data_dictionary <- read_csv("data_dictionary.csv")
lyrical_topics <- read_csv("lyrical_topics.csv")
```

-- What does it contain and what time period does it cover?

The billboard chart contains information about songs and their artists, along with data about their time on billboard charts, their genre, danceability, instruments, etc.
It covers the time period from 1958-2025.

--Which columns will be most important for your analysis?

Danceability, song structure, and weeks at number 1.

--What's missing that you might need or want?

Copies sold/streams, depending on how it was consumed at it's release. To determine general popularity besides chart numbers.

--Select a relevant column and produce basic descriptive statistics on the mean, median, lowest and highest values.

```{r}
billboard |> 
  summarize(
    mean_Danceability   = mean(Danceability, na.rm = TRUE),
    median_Danceability = median(Danceability, na.rm = TRUE),
    max_Danceability    = max(Danceability, na.rm = TRUE),
    min_Danceability    = min(Danceability, na.rm = TRUE),
    sd_Danceability = sd(Danceability, na.rm = TRUE)
  )

```

For this step, I just wanted to get a baseline of the data surrounding Danceability.

Next, propose a newsworthy hypothesis (or several of them) for your dataset. You'll be looking to find (or eliminate possible story ideas) using the techniques you've learned in this class. You'll start that effort this week by coming up with some theories or questions that you're interested in exploring that can be answered with the dataset you selected.

Null Hypothesis: Song structures do not effect the popularity of songs, nor does it effect the danceability.

Alternative Hypothesis: Specific song structures are more popular than others. Leading to more weeks at number 1. Song structure also effects danceability, with certain structures yielding higher danceability scores. 

--What questions do you have about it or need help answering?

Is there a simple way I can find streams/copies sold? Or will I have to search 1 by 1? I wasn't sure if there was a database or something.

**Part 2**

**Fix and revise your project based on the grading feedback.**

**Run at least two advanced statistical tests (t-test, Anova, z-score, etc.) that fit your project. You may need to revise the null and alternate hypothesis as necessary.**

```{r}
z_scores <- billboard |>
  transmute(

    z_weeks_number_one = (`Weeks at Number One` - mean(`Weeks at Number One`, na.rm = TRUE)) /
                          sd(`Weeks at Number One`, na.rm = TRUE),
    z_danceability     = (Danceability - mean(Danceability, na.rm = TRUE)) /
                          sd(Danceability, na.rm = TRUE)
  )


billboard |>
  select(`Weeks at Number One`, Danceability) |>
  bind_cols(z_scores) |>
  head(10)



```

For the z-scores, there were no obvious standouts. There were no scores significantly below -2 or above +2, which would be unusual. 

```{r}
anova_dance <- aov(Danceability ~ `Song Structure`, data = billboard)
summary(anova_dance)

anova_weeks <- aov(`Weeks at Number One` ~ `Song Structure`, data = billboard)
summary(anova_weeks)

```
The results of these ANOVA tests would allow up to reject the null hypothesis. The p-values are drastically below 0.05, meaning that they are statistically significant. Therefore, there is a very strong association between Song Sturctures and Danceability/popularity.

```{r}

billboard_summary <- billboard|>
  group_by(`Song Structure`) |>
  summarise(
    mean_danceability = mean(Danceability, na.rm = TRUE),
    mean_weeks_number_one = mean(`Weeks at Number One`, na.rm = TRUE),
    n_songs = n()
  ) |>
  arrange(desc(n_songs))  

billboard_summary
```

This plot was just to organize the data and see how many of each stong structure these was. Plus, it calculated the mean danceability and popularity of each structure.

**Create at least one data visualization (use ggplot)**

```{r}
structure_popularity <- billboard

ggplot(structure_popularity) +
  geom_point(aes(x = `Song Structure`, y = `Weeks at Number One`)) +
  geom_smooth(aes(x = `Song Structure`, y = `Weeks at Number One`), method = "lm") +
  expand_limits(x = 0, y = 0)

```
This graph is interesting because there appears to be a bit of a skew to the right. This isn't exactly a histogram, but it vaguely takes the shape of one. We can see that more songs in the "C" category of song structures spent more time at number one on the billboard chart.


If you encounter errors and become stuck, research the error and try to solve it as best you can. Spell out those troubleshooting steps.
