---
title: "final_project"
author: "Kaya Bogot"
date: "2025-11-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)  # For data manipulation and visualization
library(lubridate)  # For handling dates
library(knitr)      # For nice tables
library(scales)     # For better axis formatting
```
```{r}
billboard <- read_csv("billboard_hot _100_number ones_database_data.csv")
data_dictionary <- read_csv("data_dictionary.csv")
lyrical_topics <- read_csv("lyrical_topics.csv")
```

-- What does it contain and what time period does it cover?

The billboard chart contains information about songs and their artists, along with data about their time on billboard charts, their genre, danceability, instruments, etc.
It covers the time period from 1958-2025.

--Which columns will be most important for your analysis?

Danceability, song structure, and weeks at number 1.

--What's missing that you might need or want?

Copies sold/streams, depending on how it was consumed at it's release. To determine general popularity besides chart numbers.

--Select a relevant column and produce basic descriptive statistics on the mean, median, lowest and highest values.

```{r}
billboard |> 
  summarize(
    mean_Danceability   = mean(Danceability, na.rm = TRUE),
    median_Danceability = median(Danceability, na.rm = TRUE),
    max_Danceability    = max(Danceability, na.rm = TRUE),
    min_Danceability    = min(Danceability, na.rm = TRUE),
    sd_Danceability = sd(Danceability, na.rm = TRUE)
  )

```

For this step, I just wanted to get a baseline of the data surrounding Danceability.

Next, propose a newsworthy hypothesis (or several of them) for your dataset. You'll be looking to find (or eliminate possible story ideas) using the techniques you've learned in this class. You'll start that effort this week by coming up with some theories or questions that you're interested in exploring that can be answered with the dataset you selected.

Null Hypothesis: Song structures do not effect the popularity of songs, nor does it effect the danceability.

Alternative Hypothesis: Specific song structures are more popular than others. Leading to more weeks at number 1. Song structure also effects danceability, with certain structures yielding higher danceability scores. I'd also like to investigate whether certain song structures were more popular in different time periods.

--What questions do you have about it or need help answering?

Is there a simple way I can find streams/copies sold? Or will I have to search 1 by 1? I wasn't sure if there was a database or something.

**Part 2**
Fix and revise your project based on the grading feedback.

Run at least two advanced statistical tests (t-test, Anova, z-score, etc.) that fit your project. You may need to revise the null and alternate hypothesis as necessary.

```{r}
z_scores <- billboard |>
  transmute(

    z_weeks_number_one = (`Weeks at Number One` - mean(`Weeks at Number One`, na.rm = TRUE)) /
                          sd(`Weeks at Number One`, na.rm = TRUE),
    z_danceability     = (Danceability - mean(Danceability, na.rm = TRUE)) /
                          sd(Danceability, na.rm = TRUE)
  )


billboard |>
  select(`Weeks at Number One`, Danceability) |>
  bind_cols(z_scores) |>
  head(10)



```

```{r}
anova_dance <- aov(Danceability ~ `Song Structure`, data = billboard)
summary(anova_dance)

anova_weeks <- aov(`Weeks at Number One` ~ `Song Structure`, data = billboard)
summary(anova_weeks)

```
```{r}

billboard_summary <- billboard|>
  group_by(`Song Structure`) |>
  summarise(
    mean_danceability = mean(Danceability, na.rm = TRUE),
    mean_weeks_number_one = mean(`Weeks at Number One`, na.rm = TRUE),
    n_songs = n()
  ) |>
  arrange(desc(n_songs))  

billboard_summary
```


Describe and interpret the results from the statistical tests.

Create at least one data visualization (use ggplot)

```{r}
structure_popularity <- billboard

ggplot(structure_popularity) +
  geom_point(aes(x = `Song Structure`, y = `Weeks at Number One`)) +
  geom_smooth(aes(x = `Song Structure`, y = `Weeks at Number One`), method = "lm") +
  expand_limits(x = 0, y = 0)

```


If you encounter errors and become stuck, research the error and try to solve it as best you can. Spell out those troubleshooting steps.
